\chapter*{Acknowledgement}
It had been a long and sometime difficult road towards the completion of
this dissertation. I would like to express my thanks to all the people
who had made this possible. \\ \\
%
\noindent I owed my deepest gratitude to my advisor, Prof. Michael  W.
Trosset. I met Michael during an important junction of my graduate
studies when I was lost in my search for a research direction.  
Michael imparted on me the penchant for clarity of thoughts
and expression. His kindness and patience had made the process of
completing this dissertation a much more enjoyable and enlightening
experience. With high probability, I am indebted to him more than I
realized. \\ \\
%
\noindent I would like to express my thanks to my co-advisor, Prof. Dirk van
Gucht. I am grateful to Dirk for the many hours of discussion on
various topics, from data mining to uncertainty. Dirk's guidance
during the early stages of my graduate studies was crucial in
inspiring me to do research. \\ \\
%
\noindent Many thanks in particular are due to Prof. Paul W. Purdom
and Prof. David Leake. Paul's book was one of my first source in
learning about analysis of algorithms. Both Paul and David had
supervised my independent studies and I thank them for their guidance
and feedback to my semidemihemi-baked ideas. \\ \\
%
\noindent I would like to thank Dana Fielding for her words of encouragement as
well as helping me with the paperworks. In particular, Dana 
extricated me from the mess caused by my failure to signed up for the
required number of credits. I imagine it must have been as difficult
as pulling a hippopotamus from a marsh. I would also like to thank
Debbie Norris for her help in that incident. \\ \\
%
\noindent I am indebted to Prof. Carey E. Priebe for his influences in various
aspects of my work. It was the discussions between Carey and Michael
that motivated a large part of this dissertation. My work was funded
by grants where Carey and Michael are the principal investigators. \\
\\
%
\noindent I am grateful to Prof. Amol Mali and Prof. Christine Cheng from the
University of Wisconsin, Milwaukee. Amol guided me through my masters
program. Christine introduced me to combinatorics and mathematical
rigour through her lectures on graphs algorithms. \\ \\
%
\noindent I have had the pleasure of knowing a lot of nice people during my
years at IU. Special thanks to Brent Castle and Michel Salim for the
interesting and fun discussions on various topics. I thank Aaron
Jones, Bill Butske, Bledar Doraci, Carlos Castro, Darius Strapoc,
Grigor Khatchatryan, Huy Vo, Kerry Krutilla, Mikhail Gorshteyn, Paul
Kanczuzewski, Snea Thinsan, Van Ly, and Wael Abu Shammala for the
laughers and fun during the time spent chasing the soccer ball
together. \\ \\
%
\noindent Finally, I would like to thank my parents for their love and 
constant support. Without them, all my successes would be for
naught. \\ \\
%
\noindent The research described herein was supported by a grant from the Office
of Naval Research and by a subcontract to Carey E. Priebe's National
Security Science \& Engineering Faculty Fellowship.
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "dissertation"
%%% End: 
