\documentclass[12pt,reqno,final,ugsabstract,ugsabstractsigs]{iuthesis}
\pagestyle{chapter}
%\usepackage{geometry}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{mathrsfs}
\usepackage{enumerate}
\usepackage{bm}
\usepackage{tkz-graph}
\usetikzlibrary{arrows}
\usepackage{subfigure}
\usepackage[ruled]{algorithm2e}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\usepackage[colorlinks=true,pagebackref,linkcolor=magenta]{hyperref}
\usepackage[colon,sort&compress]{natbib}
\numberwithin{equation}{chapter}
\numberwithin{section}{chapter}
\renewcommand\arraystretch{1.2}
\let\underbrace\LaTeXunderbrace
\let\overbrace\LaTeXoverbrace
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
\newcommand{\argmin}{\operatornamewithlimits{argmin}}
\newcommand{\name}{Minh Tang}
\newcommand{\addr}{150 S. Woodlawn Ave, Bloomington, IN 47401}
\newcommand{\phone}{(812) 391-4892}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% New commands and environments
                
% This defines how the name looks
\newcommand{\bigname}[1]{
        \begin{center}\fontfamily{phv}\selectfont\Large\scshape#1\end{center}
}

% A ressection is a main section (<H1>Section</H1>)
\newenvironment{ressection}[1]{
	\vspace{4pt}
	{\fontfamily{phv}\selectfont\Large#1}
	\begin{itemize}
	\vspace{3pt}
}{
	\end{itemize}
}

% A resitem is a simple list element in a ressection (first level)
\newcommand{\resitem}[1]{
	\vspace{-4pt}
	\item \begin{flushleft} #1 \end{flushleft}
}

% A ressubitem is a simple list element in anything but a ressection (second level)
\newcommand{\ressubitem}[1]{
	\vspace{-1pt}
	\item \begin{flushleft} #1 \end{flushleft}
}

% A resbigitem is a complex list element for stuff like jobs and education:
%  Arg 1: Name of company or university
%  Arg 2: Location
%  Arg 3: Title and/or date range
\newcommand{\resbigitem}[3]{
	\vspace{-5pt}
	\item
	\textbf{#1}---#2 \\
	\textit{#3}
}

% This is a list that comes with a resbigitem
\newenvironment{ressubsec}[3]{
	\resbigitem{#1}{#2}{#3}
	\vspace{-2pt}
	\begin{itemize}
}{
	\end{itemize}
}

% This is a simple sublist
\newenvironment{reslist}[1]{
	\resitem{\textbf{#1}}
	\vspace{-5pt}
	\begin{itemize}
}{
	\end{itemize}
}
% \usepackage{fancyhdr}
% \pagestyle{fancy}
% \fancyhead{}
% \fancyfoot{}
% \fancyhead[RO]{\slshape \leftmark}
% \fancyfoot[C]{\thepage}
%\renewcommand{\baselinestretch}{2}
\setlength{\topmargin}{0in}
\setlength{\oddsidemargin}{0.5in}
\begin{document}
\title{Graph metrics and dimension reduction}
\author{Minh Tang}
\advisor{Michael W. Trosset}
\secondreader{Dirk van Gucht}
\thirdreader{David Leake}
\fourthreader{Paul Purdom}
\department{Informatics and Computing}
\departmentname{School}
\submitdate{October 2010}
\copyrightyear{2010}
\pagenumbering{roman}
\setcounter{page}{2}
%\input{acceptance.tex}
\begin{acknowledgements}
  It has been a long and sometimes difficult road towards the
  completion of this dissertation. I would like to express my thanks
  to all the people who has made this possible. \\ \\
%
  \noindent I owe my deepest gratitude to my advisor, Prof.~Michael
  W.  Trosset. I met Michael during an important junction of my
  graduate studies when I was lost in my search for a research
  direction.  Michael imparted on me the penchant for clarity of
  thoughts and expression. His kindness and patience had made the
  process of completing this dissertation a much more enjoyable and
  enlightening experience. With high probability, I am indebted to him
  more than I realized. \\ \\
%
  \noindent I would like to express my gratitude to my co-advisor,
  Prof.~Dirk van Gucht. I am grateful to Dirk for the many hours of
  discussion on various topics, from data mining to
  uncertainty. Dirk's guidance during the early stages of my graduate
  studies was crucial in inspiring me to do research. \\ \\
%
  \noindent I had received a lot of help and support from Prof.~Paul
  W. Purdom and Prof.~David Leake. Paul's book was one of my first
  source in learning about analysis of algorithms. Both Paul and David
  had supervised my independent studies and I thank them for their
  guidance and feedback to my semidemihemi-baked ideas. \\ \\
%
  \noindent I would like to thank Dana Fielding for her words of
  encouragement as well as helping me with the paperworks. In
  particular, Dana extricated me from the mess caused by my failure to
  signed up for the required number of credits. I imagine it must have
  been as difficult as pulling a hippopotamus from a marsh. I would
  also like to thank Debbie Norris for her help in that incident. \\ \\
%
  \noindent I am indebted to Prof.~Carey E. Priebe for his influences
  in various aspects of this work. It was the discussions between Carey
  and Michael that motivated a large part of this dissertation. My
  work was funded by grants where Carey and Michael are the principal investigators. \\
  \\
%
  \noindent I am grateful to Prof.~Amol Mali and Prof.~Christine Cheng
  from the University of Wisconsin, Milwaukee. Amol guided me through
  my masters program. Christine introduced me to combinatorics and
  mathematical rigour through her lectures on graphs algorithms. \\ \\
%
  \noindent I have had the pleasure of knowing a lot of nice people
  during my years at IU\@. Special thanks to Brent Castle and Michel
  Salim for the interesting and fun discussions on various topics. I
  thank Aaron Jones, Bill Butske, Bledar Doraci, Carlos Castro, Darius
  Strapoc, Grigor Khatchatryan, Huy Vo, Kerry Krutilla, Mikhail
  Gorshteyn, Paul Kanczuzewski, Snea Thinsan, Van Ly, and Wael Abu
  Shammala for the laughers and fun during the time spent chasing the
  soccer ball together. \\ \\
%
  \noindent Finally, I would like to thank my parents for their love
  and constant support. Without them, all my successes would be for
  naught. \\ \\
%
  \noindent The research described herein was supported by a grant
  from the Office of Naval Research and by a subcontract to Carey
  E. Priebe's National Security Science \& Engineering Faculty
  Fellowship.
\end{acknowledgements}
\begin{abstract}
  Since the introduction of Isomap and
  Locally Linear Embedding in 2000, there has
  been an explosion of interest in techniques for nonlinear dimension
  reduction.  We present a framework that unifies several prominent
  techniques, notably diffusion maps and Laplacian
  eigenmaps.  Our framework relies on the construction of various
  Euclidean distances on undirected graphs and the subsequent
  embedding of these distances in various Euclidean spaces.  We also
  consider how to construct and embed Euclidean distances on directed
  graphs.
\end{abstract}
\frontmatter
\maketitle
\signaturepage
%\copyrightpage
\makeack
\makeabstract
\tableofcontents
\mainmatter
%\input{contributions.tex}
\input{lit_review.tex}
\input{distances_undirected.tex}
\input{distances_directed.tex}
\input{dimension_reduction.tex}
\input{directed_graphs_embedding.tex}
\appendix
\input{mathematical_background}
\bibliographystyle{plainnat}
\bibliography{dissertation.bib}
\newpage
\thispagestyle{empty}
\pagestyle{empty}
%\geometry{letterpaper,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in,headheight=0in,headsep=0in,footskip=.3in}

\setlength{\parindent}{0in}
\setlength{\parskip}{0in}
\setlength{\itemsep}{0in}
\setlength{\topsep}{0in}
\setlength{\tabcolsep}{0in}

% Name and contact information



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Now for the actual document:

% Name with horizontal rule
\vspace{-1pt} {\begin{center} \Huge Curriculum Vitae \end{center}}

\bigname{\name}

\vspace{-8pt} \rule{\textwidth}{1pt}


\vspace{8 pt}

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{ressection}{Education}

	\begin{ressubsec}{Indiana University, Bloomington}{Bloomington, IN}{Ph.D in
        Computer Science}
		\ressubitem{Graduation Date: October, 2010}
	\end{ressubsec}
	\begin{ressubsec}{University of Wisconsin, Milwaukee}{Milwauke,
        WI}{M.Sc in Computer Science}
		\ressubitem{Graduation Date: May 2004}
	\end{ressubsec}
	\begin{ressubsec}{Assumption Uiversity, Thailand}{Bangkok,
        Thailand}{B.Sc in Computer Science, Summa Cum Laude}
		\ressubitem{Graduation Date: October 2001}
	\end{ressubsec}
\end{ressection}

\begin{ressection}{Research Interests}

	\resitem{Combinatorics and graph theory. Current interests include
    analytical methods for combinatorial enumeration of discrete
    structures.}

	\resitem{Statistical machine learning. Current interests include
      clustering, dimension reduction, and learning theory.}

\end{ressection}
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{ressection}{References}

	\begin{ressubsec}{Prof. Michael Trosset}{Indiana University,
        Bloomington}{Department of Statistics}
		\ressubitem{Email: mtrosset@indiana.edu}
	\end{ressubsec}
	\begin{ressubsec}{Prof. Dirk Van Gucht}{Indiana University,
        Bloomington}{School of Informatics and Computing} 
		\ressubitem{Email: vgucht@indiana.edu}
	\end{ressubsec}
	\begin{ressubsec}{Prof. Amol Mali}{University of Wisconsin,
        Milwaukee}{Department of Electrical Engineering and Computer
          Science}
		\ressubitem{Email: mali@cs.uwm.edu}
	\end{ressubsec}
	\begin{ressubsec}{Prof. Thitipong Tanprasert}{Assumption University, Thailand}{Faculty of Science and Technology}
		\ressubitem{Email: nui@scitech.au.edu}
	\end{ressubsec}
\end{ressection}


%%%%%%%%%%%%%%%%%%%%%%%%
\vfill\newpage
\end{document}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: 
