\chapter{Distances on directed graphs}
\label{cha:dist-direct-graphs}
We discussed in this chapter the extensions of some of the notions of
distances on undirected graphs in Chapter~\ref{cha:dist-undir-graphs}
to directed graphs. We concern ourselves mainly with the notions of
expected commute times, diffusion distances, and forest metrics. 
\section{Expected commute time for directed graphs}
\label{sec:expect-comm-time-1}
The notion of expected commute time as defined in
Eq. \eqref{eq:25} extends naturally onto directed graphs. Let $G =
(V,E,\omega)$ be a directed graph with similarity measure
$\omega$. Let $\mathbf{P}$ be the transition matrix of the random walk
on $G$ defined as in \S~\ref{sec:random-walks-graphs}. 
If $\mathbf{M}$ is the matrix of mean first passage time with respect
to $\mathbf{P}$, then $\mathbf{M}$ also satisfies the matrix equation
(see Proposition~\ref{prop:4}). 
\begin{equation*}
  (\mathbf{I} - \mathbf{P})\mathbf{X} = \mathbf{J} - \bm{\Pi}^{-1}
\end{equation*}
subjected to the condition 
\begin{equation*}
  \mathbf{M}_{\mathrm{dg}} = \mathbf{0}, \qquad \mathbf{M}(u,v) \geq 0   
\end{equation*}
$\mathbf{M}$ is again given by
Eq. \eqref{eq:21}. $\Delta_{\delta}$, the matrix of expected commute
time, is then given by
\begin{equation}
  \label{eq:73}
  \begin{split}
    \Delta_\delta &= \mathbf{J}(\mathbf{Z}\bm{\Pi}^{-1})_{\mathrm{dg}}
    - \mathbf{Z}\bm{\Pi}^{-1} - \bm{\Pi}^{-1}\mathbf{Z}^{T} +
    (\bm{\Pi}^{-1}\mathbf{Z})_{\mathrm{dg}}\mathbf{J} \\ \ &=
    \tfrac{1}{2}\kappa(\mathbf{Z}\bm{\Pi}^{-1} +
    \mathbf{\Pi}^{-1}\mathbf{Z}^{T})
  \end{split}
\end{equation}
% 
The following result established that $\Delta_{\delta}$ for directed
graphs is also EDM-2.
\begin{proposition}
  \label{prop:19}
  $\mathbf{Z}\bm{\Pi}^{-1} + \bm{\Pi}^{-1}\mathbf{Z}^{T}$ is
  positive definite. By Proposition~\ref{prop:18},
  $\Delta_{\delta}$ is EDM-2.  
\end{proposition}
\begin{proof}
  We use the
  following characterizations of positive definite matrices
  \citep[see][\S
  1.2]{boley09:_gener_laplac,horn94:_topic_in_matrix_analy}. If
  $\mathbf{A}$ is invertible, then
  \begin{equation}
    \label{eq:74}
    \mathbf{A} + \mathbf{A}^{T} \succ 0 \Leftrightarrow
    \mathbf{A}^{-1} + (\mathbf{A}^{-1})^{T} \succ 0
  \end{equation}
  Since $\mathbf{Z}\bm{\Pi}^{-1}$, we have
  \begin{equation}
    \label{eq:75}
    \begin{split}
      \mathbf{Z}\bm{\Pi}^{-1} + \bm{\Pi}^{-1}\mathbf{Z}^{T} \succ 0
      & \Leftrightarrow \bm{\Pi}(\mathbf{I} - \mathbf{P} + \mathbf{Q}) +
      (\mathbf{I} - \mathbf{P}^{T} + \mathbf{Q}^{T}) \bm{\Pi} \succ 0  \\
      & \Leftrightarrow \bm{\Pi}(\mathbf{I} - \mathbf{P}) +
      \bm{\Pi}(\mathbf{I} - \hat{\mathbf{P}}) + 2 \pi \pi^{T} \succ 0 \\
      & \Leftrightarrow 2 \bm{\Pi}\Bigl(\mathbf{I} - \frac{\mathbf{P} +
        \hat{\mathbf{P}}}{2}\Bigr) + 2\pi\pi^{T} \succ 0
    \end{split}
  \end{equation}
  where $\hat{\mathbf{P}}$ is the time-reversal of $\mathbf{P}$. Since
  $\hat{\mathbf{P}}$ is also a stochastic matrix,
  $\bm{\Pi}\Bigl(\mathbf{I} - \tfrac{\mathbf{P} +
    \hat{\mathbf{P}}}{2}\Bigr)$ is symmetric and diagonally dominant. Thus
  $\bm{\Pi}\Bigl(\mathbf{I} - \tfrac{\mathbf{P} +
    \hat{\mathbf{P}}}{2}\Bigr) \succeq 0$. This implies that
  $\mathbf{Z}\bm{\Pi}^{-1} + \bm{\Pi}^{-1}\mathbf{Z}^{T}$ is positive
  definite. 
\end{proof}
Some observations about expected commute time for directed graphs can
be made by representing $\mathbf{Z}\bm{\Pi}^{-1} +
\bm{\Pi}^{-1}\mathbf{Z}^{T}$ as a power series. We know from \S~\ref{sec:expect-comm-time} that $\mathbf{Z}\bm{\Pi}^{-1}$ has the
series expansion
\begin{equation*}
  \mathbf{Z}\bm{\Pi}^{-1} = \Bigl[ \mathbf{I} +
  \sum_{k=1}^{\infty}(\mathbf{P}^{k} - \mathbf{Q})\Bigr] \bm{\Pi}^{-1}
\end{equation*}
From Eq.~\eqref{eq:78} and the fact that $\mathbf{Q}\bm{\Pi}^{-1} =
\mathbf{J}$, we therefore have 
\begin{equation*}
  \begin{split}
    \bm{\Pi}^{-1} \mathbf{Z}^{T} &= \bm{\Pi}^{-1} \Bigl[ \mathbf{I} +
    \sum_{k=1}^{\infty}((\mathbf{P}^{k})^{T} - \mathbf{Q}^{T})\Bigr] \\
    &= \Bigl[ \mathbf{I} +
    \sum_{k=1}^{\infty}(\hat{\mathbf{P}}^{k} - \mathbf{Q})\Bigr]
    \bm{\Pi}^{-1}
  \end{split}
\end{equation*}
The series expansion for $\mathbf{Z}\bm{\Pi}^{-1} +
\bm{\Pi}^{-1}\mathbf{Z}^{T}$ is then given by
\begin{equation}
  \label{eq:80}
  \mathbf{Z}\bm{\Pi}^{-1} + \bm{\Pi}^{-1}\mathbf{Z}^{T} = 
  2 \Bigl[ \mathbf{I} +
  \sum_{k=1}^{\infty}\Bigl(\frac{\mathbf{P}^{k} + \hat{\mathbf{P}}^{k}}{2}
  - \mathbf{Q}\Bigr)\Bigr] \bm{\Pi}^{-1}
\end{equation}
Eq.~\eqref{eq:80} states that expected commute time between $u$ and
$v$ is composed of two parts that are basically time-reversal of each
other. This is consistent with the observation that $\mathbf{M}(u,v)$,
the mean first passage time from $u$ to $v$, can be obtained from the
time-reversal random walk starting from $v$ going back to
$u$. Eq.~\eqref{eq:80} also indicates that the symmetrization in
expected commute time for a directed graph $G$ is performed at a later
stage than that obtained by viewing $G$ as an undirected
graph. Finally, we can also view Eq.~\eqref{eq:80} as being a special
case of combining similarities from multiple graphs $G_k$, with each
graph $G_k$ having $\bm{\Pi}(\mathbf{P}^{k} + \hat{\mathbf{P}}^{k})$
as its similarity matrix.
\section{Diffusion distances for directed graphs}
\label{sec:diff-dist-direct}
The starting point for the definition of diffusion distances for
directed graphs is Eq.~\eqref{eq:43}, which we recall below.
\begin{equation}
  \label{eq:76}
  \rho^{2}_{t}(u,v) = \sum_{w \in V}{\Bigl(\mathbf{P}^{t}(u,w) -
    \mathbf{P}^{t}(v,w)\Bigr)^2 \frac{1}{\pi(w)}}
\end{equation}
The following result extends Propositions~\ref{prop:12} and
~\ref{prop:14} to directed graphs.
\begin{proposition}
  \label{prop:20}
  Let $G = (V,E,\omega)$ be a directed graph. Suppose that
  the transition matrix $\mathbf{P}$ of $G$ is irreducible. Then
  \begin{equation}
    \label{eq:72}
    \begin{split}
      \rho_{t}^{2}(u,v) &= \frac{(\mathbf{P}^{t}\hat{\mathbf{P}}^{t})(u,u) -
        (\mathbf{P}^{t}\hat{\mathbf{P}}^{t})(v,u)}{\pi(u)} +
      \frac{(\mathbf{P}^{t}\hat{\mathbf{P}}^{t})(v,v) -
        (\mathbf{P}^{t}\hat{\mathbf{P}}^{t})(u,v)}{\pi(v)}  \\
      &= (\mathbf{P}^{t}\hat{\mathbf{P}}^{t}\bm{\Pi}^{-1})(u,u) -
      (\mathbf{P}^{t}\hat{\mathbf{P}}^{t}\bm{\Pi}^{-1})(v,u) \\
      &+ (\mathbf{P}^{t}\hat{\mathbf{P}}^{t}\bm{\Pi}^{-1})(v,v) -
      (\mathbf{P}^{t}\hat{\mathbf{P}}^{t}\bm{\Pi}^{-1})(u,v)
    \end{split}
  \end{equation}
  where $\hat{\mathbf{P}}$ is the time-reversal of $\mathbf{P}$.  Thus
  $\Delta_{\rho_{t}^{2}} =
  \kappa(\mathbf{P}^{t}\hat{\mathbf{P}}^{t}\bm{\Pi}^{-1})$ is EDM-2.
\end{proposition}
\begin{proof}
  The proof of Eq.~\eqref{eq:72} is similar to that in
  Proposition~\ref{prop:12}. By the definition of time-reversal, we
  have
  \begin{equation}
    \label{eq:77}
    \pi(u) \mathbf{P}(u,v) = \pi(v) \hat{\mathbf{P}}(v,u)
  \end{equation}
  Eq.~\eqref{eq:72} can then be obtained by expanding the square in
  Eq.~\eqref{eq:76} and then using Eq.~\eqref{eq:77}. \\ \\
  % 
  \noindent We now show that
  $\mathbf{P}^{t}\hat{\mathbf{P}}^{t}\bm{\Pi}^{-1} \succeq 0$. From
  Eq.~\eqref{eq:78} we have
  \begin{equation}
    \label{eq:79}
    \mathbf{P}^{t}\hat{\mathbf{P}}^{t}\bm{\Pi}^{-1} = 
    \mathbf{P}^{t}\bm{\Pi}^{-1}(\mathbf{P}^{t})^{T}\bm{\Pi}\bm{\Pi}^{-1}
    = \mathbf{P}^{t}\bm{\Pi}^{-1}(\mathbf{P}^{t})^{T} \succeq 0
  \end{equation}
  By Proposition~\ref{prop:18}, $\Delta_{\rho_{t}^2} =
  \kappa(\mathbf{P}^{t}\hat{\mathbf{P}}^{t}\bm{\Pi}^{-1})$ is EDM-2.
\end{proof}
The relationship between diffusion distances and expected commute time
as stated in Proposition~\ref{prop:17} for the case when $G$ is an
undirected graph fails to hold when $G$ is a directed graph. This is
because diffusion distance for directed graphs is based on
$\mathbf{P}^{t}\hat{\mathbf{P}}^{t}$ while expected commute time is
based on $(\mathbf{P}^{t} + \hat{\mathbf{P}}^{t})/2$ (see
Eq.~\eqref{eq:80}) and the two are not as closely related as they
would have been had $G$ been an undirected graph. 
\section{Forest metrics for directed graphs}
\label{sec:forest-metr-direct}
We now discuss forest metrics
\citep{chebotarev02:_fores_metric_for_graph_vertic} for directed
graphs. If we extends forest metrics to directed graphs by using
Eq.~\eqref{eq:30} and Eq.~\eqref{eq:41}, then the resulting distance
might not be Euclidean distances. Consider for example the graph in
Figure~\ref{fig:directed_graph1}. The matrix $\Delta =
\kappa((\mathbf{I} + 4 \mathbf{L})^{-1})$ is not EDM-2.
\begin{figure}[hbtp]
  \centering
  \begin{minipage}[c]{0.45\textwidth}
    \centering
    \begin{tikzpicture}[node distance = 1.5 cm]
      \SetUpEdge[lw  = 1.5pt, color = orange, labelcolor = red!30, 
      labelstyle = {draw,sloped}]
      \GraphInit[vstyle = Normal]
      \tikzset{VertexStyle/.append style = {fill = red!50}}
      \Vertex[x = 0, y = 0]{5}
      \Vertex[x = 3, y = 0]{4}
      \Vertex[x = 3, y = 3]{2}
      \Vertex[x = 0, y = 3]{1}
      \Vertex[x = 5, y = 1.5]{3}
      \tikzset{EdgeStyle/.style={post}}
      \Edge[label=$4$](5)(1)
      \Edge[label=$5$](4)(2)
      \Edge[label=$2$](4)(5)
      \Edge[label=$20$](2)(1)
      \Edge[label=$50$](2)(3)
      \Edge[label=$4$](3)(4)
      \Edge[label=$1$](1)(4)
    \end{tikzpicture}
  \end{minipage}
  \begin{minipage}[c]{0.38\textwidth}
    \begin{equation*}
      \mathbf{L} = \left[ \begin{array}{rrrrr}
          1  & 0  & 0  & -1  & 0 \\
          -20 & 70  & -50  & 0 & 0 \\
          0  & 0 & 4  & -4 & 0 \\
          0 & -5 & 0  & 7 & -2 \\
          -4 & 0 & 0 & 0 & 4
        \end{array} \right ]
    \end{equation*}
  \end{minipage}
  \caption{A directed graph and its Laplacian matrix based on outdegrees.}
  \label{fig:directed_graph1}
\end{figure}
Another possible extension of forest metrics is by using the
relationship between forest metrics and expected commute time as
stated in Proposition~\ref{prop:9}. Specifically, let $G =
(V,E,\omega)$ be a directed graph. Augment a vertex $v_{*}$ to $G$ and
connect $v_{*}$ to all vertices of $G$ with edge weight $1$, with the
connection being undirected. Compute the expected commute time on
$\delta_{G_{*}}$ on $G_*$ and scale $\delta_{G_{*}}$ by a constant to
get the forest metrics on $G$. This extension of forest metrics
satisfies the property that the resulting distance is an Euclidean
distance since expected commute time was shown to be an Euclidean
distance for general graphs in \S~\ref{sec:expect-comm-time-1}. A
possible downside to this extension of forest metrics is that the
resulting distance is just expected commute time on some
different graph. \\ \\
% 
% 
\noindent 
We now discuss a possible extension of forest metrics that's different
from the two approaches mentioned above. Our approach leads to an
Euclidean distance that doesn't seem to have an obvious relationship
with expected commute time. Let $G = (V,E,\omega)$ be a graph with
similarity measure $\omega$. Let $\mathbf{P}$ be the transition matrix
of $G$. We assumed that $\mathbf{P}$ is irreducible. Consider the
matrix $\mathbf{X}_{\beta} = \mathbf{I} + \alpha \bm{\Pi}(\mathbf{I} -
\mathbf{P})$ with $\beta \geq 0$. Note that for an undirected graph
$G$, $\mathbf{L} = \mathrm{Vol}(G) \bm{\Pi}(\mathbf{I} -
\mathbf{P})$. Thus $\mathbf{X}_{\beta}$ subsumes the role of
$(\mathbf{I} + \alpha \mathbf{L})$ for general graphs. \\ \\
% 
% 
Define the notion of forest metrics for directed graph as
\begin{equation}
  \label{eq:81}
  \eta_{\beta}(u,v) = \mathbf{X}_\beta^{-1}(u,u) - \mathbf{X}_\beta^{-1}(u,v) -
  \mathbf{X}_\beta^{-1}(v,u) + \mathbf{X}_\beta^{-1}(v,v) 
\end{equation}
That is, $\Delta_{\eta_{\beta}} = \kappa(\mathbf{X}_\beta^{-1}) =
\tfrac{1}{2}\kappa(\mathbf{X}_\beta^{-1} +
(\mathbf{X}_{\beta}^{-1})^{T})$. The next result shows that
$\Delta_{\eta_{\beta}}$ as defined is EDM-2.
\begin{proposition}
  \label{prop:23}
  Let $\mathbf{X}_{\beta}$ be defined as above. Then
  $\mathbf{X}_{\beta}^{-1} + (\mathbf{X}_{\beta}^{-1})^{T}$ is
  positive definite. $\Delta_{\eta_{\beta}}$ as defined in
  Eq.~\eqref{eq:81} is EDM-2.
\end{proposition}
\begin{proof}
  The proof is similar to that of Proposition~\ref{prop:19}.  We use
  the characterization of positive definite matrix as in
  Eq.~\eqref{eq:74}, i.e.,
  \begin{equation*}
    \mathbf{A} + \mathbf{A}^{T} \succ 0 \Leftrightarrow
    \mathbf{A}^{-1} + (\mathbf{A}^{-1})^{T} \succ 0
  \end{equation*}
  First of all, $\mathbf{X}_{\beta}$ is strictly diagonally dominant
  and hence invertible. Secondly, we have $\mathbf{X}_{\beta}^{T} =
  \mathbf{I} + \alpha (\mathbf{I} - \mathbf{P}^{T}) \bm{\Pi} =
  \mathbf{I} + \alpha \bm{\Pi} (\mathbf{I} - \hat{\mathbf{P}})$ where
  $\hat{\mathbf{P}}$ is the time-reversal of $\mathbf{P}$. Therefore, 
  \begin{equation}
    \label{eq:82}
    \mathbf{X}_{\beta} + \mathbf{X}_{\beta}^{T} = 2 \mathbf{I}
    + 2 \bm{\Pi}\Bigl(\mathbf{I} - \frac{\mathbf{P} + \hat{\mathbf{P}}}{2}
    \Bigr)
  \end{equation}
  Now, $\mathbf{X_{\beta}} + (\mathbf{X}_{\beta})^{T}$ is symmetric and
  strictly diagonally dominant. $\mathbf{X}_\beta +
  \mathbf{X}_{\beta}^{T}$ is therefore positive definite as claimed.  
\end{proof}
It's still the case that $\mathbf{X}_{\beta}^{-1}$ is a stochastic
matrix like $Q_{\alpha}$ in \S~\ref{sec:forest-metrics}. This can be seen as
follow. $\mathbf{X}_{\beta}$ is a strictly diagonally dominant matrix
with non-positive off-diagonal entries and hence is a $M$-matrix (see
\S~\ref{sec:matrix-analysis}). By
Theorem~\ref{thm:2}, $\mathbf{X}_{\beta}^{-1}$ exists and is
nonnegative. The vector $\bm{1}$ of all ones is an eigenvector of
$\mathbf{X}_{\beta}$ with eigenvalue $1$ since $(\mathbf{I} -
\mathbf{P})\bm{1} = 0$ and therefore $\bm{1}$ is also an eigenvector
of $\mathbf{X}_{\beta}^{-1}$ with eigenvalue $1$. Since
$\mathbf{X}_{\beta}^{-1}$ is a non-negative matrix, we thus have that
$\mathbf{X}_{\beta}^{-1}$ is a stochastic
matrix. \\ \\
% We summarized the above observations in the following
% proposition.
% \begin{proposition}
%   \label{prop:21}
%   Let $\mathbf{X}_{\beta} = \mathbf{I} + \beta \bm{\Pi}(\mathbf{I} -
%   \mathbf{P})$ for some fixed $\beta \geq 0$. Then
%   $\mathbf{X}_{\beta}^{-1}$ exists and is a stochastic
%   matrix. 
% \end{proposition}
% 
% 
Now, if $G$ is an undirected graph, $\mathbf{X}_{\beta} =
\mathbf{X}_{\beta}^{T}$, and $\mathbf{X}_{\beta}^{-1} =
(\mathbf{X}_{\beta}^{-1})^{T} = \mathbf{Q}_{\alpha}$ for $\alpha =
\beta/\mathrm{Vol}(G)$. Proposition~\ref{prop:23} generalizes
Theorem~\ref{thm:4} to general graphs as desired.

\section{$f(\mathbf{P} - \mathbf{Q})$ and graph metrics} \label{sec:fmathbfp-mathbfq-gra} We saw in
\S~\ref{sec:graph-metr-funct} that a general notion of Euclidean distances on graphs can be defined in terms of matrix
functions acting on $\mathbf{P} - \mathbf{Q}$. It would be desirable if the claims in Proposition~\ref{prop:13} were to
also hold for directed graphs. However, this is not the case. Consider for example the matrix function $f = 1/(1-x)^{k}$
for some integer $k \geq 1$. Proposition~\ref{prop:13} established that $\kappa(f(\mathbf{P} -
\mathbf{Q})\bm{\Pi}^{-1})$ is an Euclidean distance for any $k \geq 1$.  The following result showed that this is false
when the underlying graph $G$ is directed.  \begin{proposition} \label{prop:25} Let $G = (V,E,\omega)$ be a directed
  graph and $\mathbf{P}$ be the transition matrix on $G$. Let $f_k = 1/(1-x)^{k}$ for integer $k \geq 1$. Then there
  exists an integer $n_0 \geq 2$ such that the symmetric part of $f_{n_0}(\mathbf{P} - \mathbf{Q})\bm{\Pi}^{-1}$ is
  indefinite.  \end{proposition} \begin{proof} Theorem 1 from \citet{johnson75:_power_matric_posit_defin_real_part}
  stated that if a matrix $\mathbf{A}$ has positive semidefinite Hermitean part $H(\mathbf{A})$, then
  $H(\mathbf{A}^{m}) \succeq 0$ for all integer $m \geq 1$ if and only if $\mathbf{A}$ is Hermitean. We know from
  Proposition~\ref{prop:19} that $H(f_1(\mathbf{P} - \mathbf{Q})\bm{\Pi}^{-1}) = H((\mathbf{I} - \mathbf{P} +
  \mathbf{Q})^{-1}\bm{\Pi}^{-1})$ is positive semidefinite. Note that $H(f_{k}(\mathbf{P} - \mathbf{Q})\bm{\Pi}^{-1})$
  is positive semidefinite if and only if $H(\bm{\Pi}^{1/2}f_{k}(\mathbf{P} - \mathbf{Q}) \bm{\Pi}^{-1}
  \bm{\Pi}^{1/2})$ is positive semidefinite. Thus $H(f_{k}(\mathbf{P} - \mathbf{Q})\bm{\Pi}^{-1}) \succeq 0$ if and
  only if $H((\bm{\Pi}^{1/2}f_{1}(\mathbf{P} - \mathbf{Q})\bm{\Pi}^{-1/2})^{k}) \succeq 0$. Therefore,
  $H(f_{k}(\mathbf{P} - \mathbf{Q})\bm{\Pi}^{-1}) \succeq 0$ for all $k \geq 1$ if and only if
  $\bm{\Pi}^{1/2}f_{1}(\mathbf{P} - \mathbf{Q})\bm{\Pi}^{-1/2}$ is Hermitean, which is not the case. Therefore, there
  exists an integer $n_0 \geq 2$ such that $H(f_{n_0}(\mathbf{P} - \mathbf{Q})\bm{\Pi}^{-1})$ is indefinite as
  claimed.  \end{proof}
% 
% 
Even though Proposition~\ref{prop:13}, and similarly,
Proposition~\ref{prop:24}, failed to hold for directed graphs, we
can establish a much weaker result in the same vein for relaxed
random walks.
\begin{proposition}
  \label{prop:26}
  Let $G$ be a directed graph and $\mathbf{P}$ be the transition
  matrix on $G$. Suppose that $\mathbf{P}$ is irreducible and
  aperiodic. Let $f(x)$ be a function with a series expansion $f(x) =
  \sum_{k \geq 0}{a_k x^k}$ with $a_0 > 0$, $a_k \geq 0\,\,\, \forall
  k \geq 1$. Suppose that the radius of converge of $f$ is $R \geq
  1$. Then there exist a $\alpha \in (0,1)$ such that
  $H(f(\mathbf{P}_\alpha - \mathbf{Q}) \bm{\Pi}^{-1})$ is a positive
  semidefinite matrix, where $\mathbf{P}_\alpha = (1 -
  \alpha)\mathbf{I} + \alpha \mathbf{P}$. $\kappa( f(\mathbf{P}_\alpha
  - \mathbf{Q})\bm{\Pi}^{-1})$ is then an EDM-2
  matrix.  
\end{proposition}
%
\begin{proof}
  For any $\alpha \in (0,1)$, $\mathbf{P}_\alpha$ is the transition
  matrix for a relaxed random walk on $G$.  Furthermore, the
  stationary distribution of $\mathbf{P}$ and $\mathbf{P}_{\alpha}$
  coincides, and $\lim_{k \rightarrow \infty} \mathbf{P}_{\alpha}^{k}
  = \mathbf{Q} = \mathbf{1}\bm{\pi}^{T}$. To establish the claim in
  Proposition~\ref{prop:26}, it is sufficient to show that, for any
  irreducible and aperiodic $\mathbf{P}$, there exist a $\alpha \in
  (0,1)$ such that $\mathbf{P}_\alpha - \mathbf{Q}$ is diagonally
  dominant. Assume that we can show that for any irreducible and
  aperiodic $\mathbf{P}$, there exist a $\alpha \in (0,1)$ such that
  $\mathbf{P}_{\alpha} - \mathbf{Q}$ is diagonally dominant. Then, for
  any $\beta \in (0,\alpha)$, $\mathbf{P}_{\beta} - \mathbf{Q} =
  \mathbf{P}_{\alpha} - \mathbf{Q} + (\alpha - \beta)(\mathbf{I} -
  \mathbf{P})$ is also diagonally dominant. It also follows that for
  any integer $k \geq 1$, there exists some $\alpha_k \in (0,1)$ such
  that $((1 - \alpha_k)\mathbf{I} + \alpha_k\mathbf{P})^{k} -
  \mathbf{Q}$ is diagonally dominant. By taking $\alpha_* =
  \min\{\alpha_1, \alpha_2, \dots, \alpha_M\}$, we can show that for
  any integer $M \geq 1$, there exists some $\alpha_* \in (0,1)$ such
  that $((1 - \alpha_*)\mathbf{I} + \alpha_* \mathbf{P})^{k} -
  \mathbf{Q}$ is diagonally dominant for $1 \leq k \leq M$. For $l
  \geq M + 1$, we can use the following bound on the rate of
  convergence to the stationary distrpibution. If $\mathbf{P}$ is
  irreducible and aperiodic and $\bm{\pi}$ is the stationary
  distribution of $\mathbf{P}$, then there exist constants $C$ and
  $\alpha \in (0,1)$ such that
  \begin{equation} 
    \label{eq:118}
    \max_{x} \| \mathbf{P}^{k}(x,\cdot) - \bm{\pi} \|_{TV} \leq C
    \alpha^{k} 
  \end{equation}
  where $\| \cdot \|_{TV}$ is the total variation distance. Using
  Eq.~(\ref{eq:118}), one can show that for any $a_0 > 0$, there
  exists an integer $M \geq 1$ such
  that 
  \begin{equation} 
    \label{eq:119} 
    a_0 \mathbf{I} + \sum_{l = M +
      1}^{\infty} a_k[ (\mathbf{P}_{\alpha_*})^{k} -
    \mathbf{Q}] 
  \end{equation}
  is diagonally dominant, provided that $\limsup_{k \rightarrow
    \infty}{|a_k|^{1/k}} \leq 1$.  $f(\mathbf{P}_{\alpha} -
  \mathbf{Q})$ is thus diagonally dominant for some $\alpha \in (0,1)$
  because the $a_k$ are all non-negative and the radius of convergence
  of $f$ is at least $1$. Because $H(f(\mathbf{P}_{\alpha} -
  \mathbf{Q})\bm{\Pi}^{-1}) = f(\mathbf{P}_{\alpha} -
  \mathbf{Q})\bm{\Pi}^{-1} + f(\hat{\mathbf{P}}_{\alpha} -
  \mathbf{Q})\bm{\Pi}^{-1}$ where $\hat{\mathbf{P}}_{\alpha}$ is the
  time-reversal of $\mathbf{P}_{\alpha}$,
  $H(f(\mathbf{P}_{\alpha} - \mathbf{Q})\bm{\Pi}^{-1})$ is diagonally dominant for some $\alpha \in (0,1)$. \\ \\
  % 
  \noindent To show that $\mathbf{P}_{\alpha} - \mathbf{Q}$ is
  diagonally dominant for some $\alpha \in (0,1)$, one must show
  that there exist a $\alpha \in (0,1)$ such
  that 
  \begin{equation} 
    \label{eq:115} 
    (1 - \alpha) + \alpha
    \mathbf{P}(i,i) - \pi_i \geq \sum_{j}|\alpha \mathbf{P}(i,j) -
    \pi_j| \quad \forall i 
  \end{equation}
  The inequality in Eq.~\eqref{eq:115} is certainly true for any
  $\alpha$ such that
  \begin{equation}
    \label{eq:116} \alpha \leq \pi_i \leq 1 - \alpha \quad \forall i 
  \end{equation} 
  Because $\mathbf{P}$ is irreducible, $\pi_i > 0$ for all $i$ and a
  $\alpha$ that satisfy Eq.~\eqref{eq:116} is 
  \begin{equation}
    \label{eq:117} \alpha = \min( \min_{i}\{ \pi_i \}, \max_{i} \{1 -
    \pi_i \}) 
  \end{equation} 
  The matrix $\mathbf{P}_{\alpha} - \mathbf{Q}$ is thus diagonally
  dominant for some $\alpha \in (0,1)$.  
\end{proof} 
Note that Proposition~\ref{prop:26} only indicated that for any
transition matrix $\mathbf{P}$ and function $f$ with non-negative
coefficients, there exist a $\alpha \in (0,1)$ such that
$H(f(\mathbf{P}_{\alpha} - \mathbf{Q})\bm{\Pi}^{-1})$, i.e., the
$\alpha$ is dependent on $\mathbf{P}$ and $f$. Contrast this with
Proposition~\ref{prop:24} where $\alpha$ is independent of
$\mathbf{P}$. The dependence of $\alpha$ on $\mathbf{P}$ decreases the
usefulness of Proposition~\ref{prop:26} significantly. However, there
are scenarios in which one can find a $\alpha$ that is independent of
$\mathbf{P}$. For example, the matrix $\exp((\mathbf{I} +
\mathbf{P})/2)\bm{\Pi}^{-1}$ has positive semidefinite Hermitean
part. This can be seen as follows. The matrix $\exp((\mathbf{I} +
\mathbf{P})/2$ has row sum $\exp(1)$. Because the diagonal entries of
$\exp((\mathbf{I} + \mathbf{P})/2)$ are at least $\tfrac{3}{2}$, the
matrix $\exp((\mathbf{I} + \mathbf{P})/2)$ is diagonally
dominant. $H(\exp((\mathbf{I} + \mathbf{P})/2)\bm{\Pi}^{-1}) =
\exp((\mathbf{I} + \mathbf{P})/2)\bm{\Pi}^{-1} + \exp((\mathbf{I} +
\hat{\mathbf{P}})/2)\bm{\Pi}^{-1}$ is thus diagonally dominant and
hence positive semidefinite.  $\exp((\mathbf{I} +
\mathbf{P})/2)\bm{\Pi}^{-1}$ can be viewed as a diffusion kernel on
directed graphs, extending \citet{kondor02:_diffus} notion of
diffusion kernels for undirected graphs.
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "dissertation"
%%% End: 
% We use the following characterization of positive definite matrix. Let $\mathbf{A}$ be invertible. Then $\mathbf{A} +
% \mathbf{A}^{T}$ is positive definite if and only if $\mathbf{A}^{-1} + (\mathbf{A}^{-1})^{T}$ is positive definite
% \citep[see][\S 1.2]{horn94:_topic_in_matrix_analy}. Since $\mathbf{Z}\bm{\Pi}^{-1}$ is invertible, \begin{equation}
%   \label{eq:34} \begin{split} \mathbf{Z}\bm{\Pi}^{-1} + \bm{\Pi}^{-1}\mathbf{Z}^{T} \succ 0 & \leftrightarrow
%     \bm{\Pi}(\mathbf{I} - \mathbf{P} + \mathbf{Q}) + (\mathbf{I} - \mathbf{P}^{T} + \mathbf{Q}^{T})\bm{\Pi} \succ 0 \\ &
%     \leftrightarrow \bm{\Pi}(\mathbf{I} - \mathbf{P}) + (\mathbf{I} - \mathbf{P}^{T})\bm{\Pi} + 2 \pi \pi^{T} \succ 0 \\ &
%     \leftrightarrow \bm{\Pi}(\mathbf{I} - \frac{\mathbf{P} + \mathbf{P}_{*}}{2}) + 2 \pi \pi^{T} \succ 0 \end{split}
% \end{equation} where $\mathbf{P}_{*}$ is the time-reverseral of $\mathbf{P}$. Now, $\bm{\Pi}(\mathbf{I} -
% \frac{\mathbf{P} + \mathbf{P}_{*}}{2})$ is symmetric. Furthermore, $\frac{\mathbf{P} + \mathbf{P}_{*}}{2}$ is a
% stochastic matrix and thus $\bm{\Pi}(\mathbf{I} - \frac{\mathbf{P} + \mathbf{P}_{*}}{2})$ is diagonally dominant. By
% Ger\u{s}gorin's circle theorem \citep{gersgorin31:_uber_abgren_eigen_matrix}, $\bm{\Pi}(\mathbf{I} - \frac{\mathbf{P}
% + \mathbf{P}_{*}}{2})$ is positive semidefinite. Thus, $\mathbf{Z}\bm{\Pi}^{-1} + \bm{\Pi}^{-1}\mathbf{Z}^{T} \succ
% 0$, and $\Delta_{\delta}$ is an EDM-2 matrix.  Forest metrics
% An interpretation for the entries in $\mathbf{Q}_{\alpha}$ is also available in
% \citep{chebotarev02:_fores_metric_for_graph_vertic}. Specifically, let 
