\documentclass{article}
\usepackage{mystyle}
\begin{document}
\title{Dissertation Outline}
\author{Minh Tang}
\maketitle
\begin{enumerate}
\item Introduction and discussion about the scope of the
  thesis. Scope of the thesis is
  \begin{itemize}
  \item Notions of distances on directed and undirected graphs
    based on random walks on the graphs.
  \item Relationships between different notions of graph distances.
  \item Applications of graph distances to manifold learning and
    dimensionality reduction.
  \item[*] Other possible applications of graph distances,
    e.g. clustering, classification.
  \end{itemize}
\item Representations of data. Discuss the representation of data as
  \begin{itemize}
  \item Features vector / Features space.
  \item Proximity measure.
  \item Graphs.
  \end{itemize}
  Discuss the interplay between the representations. Mention briefly
  distance geometry, classical MDS, graph Laplacians, and random walks
  on graphs.
\item Literature review. We will mentioned the relevant literature
  related to the following topics.
  \begin{itemize}
  \item Manifold learning algorithms, e.g., Laplacian eigenmaps, LLE,
    Hessian eigenmaps, diffusion maps, LTSP.
  \item Distances on graphs, e.g., expected commute time, resistance
    distance, forest metrics, diffusion distances, diffusion kernel.
  \item Application of graphs distances, e.g., classification,
    clustering, semi-supervised learning (if any).
  \item[*] Other topics, e.g. kernel regularization of graph
    Laplacian, graph distances as number of points increased to infinity.
  \end{itemize}
\item Distances on undirected graph and their relationships.
  \begin{itemize}
  \item Expected commute time.
  \item Diffusion distance.
  \item Forest metrics.
  \item[*] A general class of distance measure of the form
    \begin{equation}
      \label{eq:1}
      f(\mathbf{P} - \mathbf{Q}) = \sum_{k = 0}^{\infty}a_k(\mathbf{P}- \mathbf{Q})^{k}
    \end{equation}
    where $f$ is analytic inside the unit ball and $f((-1,1)) \subset
    \mathbb{R}^{\geq 0}$. 
  \end{itemize}
\item Distances on directed graphs.
\item Graph distances, manifold learning, and dimensionality
  reduction. 
  \begin{itemize}
  \item Relationship between graph distances and embedding using
    classical MDS. Canonical representation. Kernel view.
  \item  Embeddings obtained from Laplacian eigenmaps vs diffusion maps.
  \item[*] LLE, extension of Belkin and Niyogi Laplacian eigenmaps to
    directed graphs.
  \end{itemize}
\item Some experimental result and data analysis.
\item Conclusion and future work.
\end{enumerate}
\end{document}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: 
